---
title: "Hurricane_2011-2016"
author: "Griffin Dean"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#important question: This study will focus on the following research question: 
#how has the U.S. Virgin Islands’ inter-seasonal weather patterns from 2000-2022
indicate a deviation from traditional hurricane season?

## Questions asked
-when did the hurricanes occur, 
-compare with hurricane season in the virgin islands, 
-if it is outside the hurricane season, how much does it devaite

#large timespan with no hurricanes, then many different hurricanes in a short period of time

#ocean data pull tools
 #if hurricanes come in at sunrise or sunspot.year

#peak windspeed in comparison to mean windspeed in a yea

#plot wind speeds in a year vs wind speed vs a hurricane
```{r}
library(tidyverse)
erika <- read.csv(file="erika.csv", header=TRUE, sep=",")
rafael <- read.csv(file="rafael.csv", header=TRUE, sep=",")
fullData <- read.csv(file="GriffDeanData.csv", header=TRUE, sep=",")
```

## Hurricane Erika data(august 2015)
```{r Hurricane Erika}
erika.knots <- erika$knots
erika.knots.mean <- mean(erika.knots, na.rm=FALSE)
erika.knots.SD <- sd(erika.knots)
print("Mean windspeed erika: ", erika.knots.mean )
print(erika.knots.mean)

erika.windDir <- erika$windDir
erika.windDir.mean <- mean(erika.windDir, na.rm=FALSE)
erika.windDir.SD <- sd(erika.windDir)
print("Mean wind direction erika:" )
print(erika.windDir.mean)

erika.gustKnots <- erika$gustKnots
erika.gustKnots.mean <- mean(erika.gustKnots, na.rm=FALSE)
erika.gustKnots.SD <- sd(erika.gustKnots)
print("Mean gustKnots erika:" )
print(erika.gustKnots.mean)

erika.airTemp <- erika$airTemp
erika.airTemp.mean <- mean(erika.airTemp, na.rm=FALSE)
erika.airTemp.SD <- sd(erika.airTemp)
print("Mean airtemp erika:" )
print(erika.airTemp.mean)

erika.pressure <- erika$pressure
erika.pressure.mean <- mean(erika.pressure, na.rm=FALSE)
erika.pressure.SD <- sd(erika.pressure)
print("Mean pressure erika:" )
print(erika.pressure.mean)

erika.waterTemp <- erika$waterTemp
erika.waterTemp.mean <- mean(erika.waterTemp, na.rm=FALSE)
erika.waterTemp.SD <- sd(erika.waterTemp)
print("Mean waterTemp erika:" )
print(erika.waterTemp.mean)
```

## Hurricane Rafael(oct 2012)
-Data Processing
```{r Hurricane Rafael}
rafael.knots <- rafael$knots
rafael.knots.mean <- mean(rafael.knots, na.rm=FALSE)
rafael.knots.SD <- sd(rafael.knots)
print("Mean windspeed rafael:" )
print(rafael.knots.mean)

rafael.windDir <- rafael$windDir
rafael.windDir.mean <- mean(rafael.windDir, na.rm=FALSE)
rafael.windDir.SD <- sd(rafael.windDir)
print("Mean wind direction rafael:" )
print(rafael.windDir.mean)

rafael.gustKnots <- rafael$gustKnots
rafael.gustKnots.mean <- mean(rafael.gustKnots, na.rm=FALSE)
rafael.gustKnots.SD <- sd(rafael.gustKnots)
print("Mean gust rafael:" )
print(rafael.gustKnots.mean)

rafael.airTemp <- rafael$airTemp
rafael.airTemp.mean <- mean(rafael.airTemp, na.rm=FALSE)
rafael.airTemp.SD <- sd(rafael.airTemp)
print("Mean airtemp rafael:" )
print(rafael.airTemp.mean)

rafael.pressure <- rafael$pressure
rafael.pressure.mean <- mean(rafael.pressure, na.rm=FALSE)
rafael.pressure.SD <- sd(rafael.pressure)
print("Mean pressure rafael:" )
print(rafael.pressure.mean)

rafael.waterTemp <- rafael$waterTemp
rafael.waterTemp.mean <- mean(rafael.waterTemp, na.rm=FALSE)
rafael.waterTemp.SD <- sd(rafael.waterTemp)
print("Mean waterTemp rafael:" )
print(rafael.waterTemp.mean)
```

## Full data set
-2011
-2016
```{r All Data}
fullData.knots <- fullData$knots
fullData.knots.mean <- mean(fullData.knots, na.rm=FALSE)
fullData.knots.SD <- sd(fullData.knots)
print("Mean windspeed fulldata:" )
print(fullData.knots.mean)

fullData.windDir <- fullData$windDir
fullData.windDir.mean <- mean(fullData.windDir, na.rm=FALSE)
fullData.windDir.SD <- sd(fullData.windDir)
print("Mean windDir fulldata:" )
print(fullData.windDir.mean)

fullData.gustKnots <- fullData$gustKnots
fullData.gustKnots.mean <- mean(fullData.gustKnots, na.rm=FALSE)
fullData.gustKnots.SD <- sd(fullData.gustKnots)
print("Mean gust Fulldata:" )
print(fullData.gustKnots.mean)

fullData.airTemp <- fullData$airTemp
fullData.airTemp.mean <- mean(fullData.airTemp, na.rm=FALSE)
fullData.airTemp.SD <- sd(fullData.airTemp)
print("Mean temp fullData:" )
print(fullData.airTemp.mean)

fullData.pressure <- fullData$pressure
fullData.pressure.mean <- mean(fullData.pressure, na.rm=FALSE)
fullData.pressure.SD <- sd(fullData.pressure)
print("Mean pressure fulldata:" )
print(fullData.pressure.mean)

fullData.waterTemp <- fullData$waterTemp
fullData.waterTemp.mean <- mean(fullData.waterTemp, na.rm=FALSE)
fullData.waterTemp.SD <- sd(fullData.waterTemp)
print("Mean waterTemp fulldata:" )
print(fullData.waterTemp.mean)
```

#Summary
```{r}
summary(erika)
summary(rafael)
summary(fullData)
```
```{r}
library(ggplot2)
ggplot.Erica <- ggplot(erika, aes(pressure, airTemp))
ggplot.Rafael <- ggplot(rafael, aes(pressure, airTemp))
ggplot.All_Time <- ggplot(fullData, aes(pressure, airTemp))
```

```{r}
a <- ggplot.Erica
a + geom_label(aes(label = Erica), nudge_x = 1,nudge_y = 1) 

a + geom_boxplot()

```
