---
title: "Hypothesis Testing"
output: html_document
---

## Section1: Z-Test
> Lower Tail Test of Population Mean with Known Variance lwrTailMean.R

```{r}
mu0 <- 10000    # hypothesized value
xbar <- 9900    # sample mean
sigma <- 120    # population standard deviation
n <- 30         # sample size

z <- (xbar-mu0)/(sigma/sqrt(n)) # test statistic

print("z is equal to")
print(z)

# Calculate the critical value
alpha <- 0.05
z.alpha <- qnorm(1-alpha)

print("critical value is equal to")
print(-z.alpha)

if(z <= z.alpha){
  print("Reject the claim that mean lifetime of a light bulb is above 10,000 hours")
} else {
  print("Do not reject the claim that mean lifetime of a light bulb is above 10,000 hours")
}
```

### Section 1:Uppertail

> Upper Tail Test of Population Mean with Known Variance Upertail.R

```{r}
muO <- 2        # hypothesized value
xbar <- 2.1     # sample mean
sigma <- 0.25   # population standard deviation
n <- 35         # sample size

z <- (xbar-muO)/(sigma/sqrt(n))
print("z is equal to ")
print(z)

#compute the critical value at.05 significance level
alpha <- 0.5
z.alpha <- qnorm(1-alpha)

print("z.alpha is equal to")
print(z.alpha)

if(z>=z.alpha){
  print("We reject the claim that there is at most 2 grams of saturated fat in a cookie")
} else {
  print("Do not reject the claim that there is at most 2 grams of saturated fat in a cookie")
}
```

### Section 1: Bothtail

> Two-Tailed Test of Population Mean with Known Variance Bothtail.R

```{r}
muO <- 15.4     # hypothesized value
xbar <- 14.6    # sample mean
sigma <- 2.5    # population standard deviation
n <- 35         # sample size

z <- (xbar-muO)/(sigma/sqrt(n))
print("z is equal to ")
print(z)

#compute the critical value at.05 significance level
alpha <- 0.5
z.alpha <- qnorm(1-alpha)

print("z.alpha is equal to")
print(z.alpha)

if(z>=z.alpha){
  print("We reject the claim that there is at most 2 grams of saturated fat in a cookie")
} else {
  print("Do not reject the claim that there is at most 2 grams of saturated fat in a cookie")
}
```

### Section 1: Z-Test

> zVal-ReadFile.R ztest-a.csv

```{r}
# sample mean: $100,200
# population sd: $15,000
# standard error: $2,739
# sample size: 30
# z-score: -4.67
# HO (GlassDoor data): $113,000
# two-sided test

sal <- read.csv("ztest-a.csv")
# Get the sample mean from the summary of the file
print(summary(sal))

#Create a function that calculates z test
z.test <- function(a, mu, sd){
  zeta <- (mean(a) - mu)/(sd/sqrt(length(a)))  # n = length(a)
  return(zeta)
}

z <- z.test(a = sal$salary, mu = 113000, sd = 15000)
print("z value is equal to: ")
print(z)

alpha <- 0.05
z.half.alpha <- qnorm(1-alpha/2)
print("-z.half.alpha, z.half.alpha is equal:")
print(c(-z.half.alpha, z.half.alpha))

if(z <= -z.half.alpha || z >= z.half.alpha){
  print("reject the null hypothesis")
} else {
  print("do not reject the null hypothesis")
}
```

## Section 2: t-Test

### Lower Tail Test of Population Mean with Unknown Variance

> tTestLwr.R

```{r}
muO <- 2          # hypothesized value
xbar <- 1.8       # sample mean
n <- 10           # population standard deviation
s <- 0.15         # sample size

t <- (xbar-muO)/(s/sqrt(n))
print("The value of t is equal to ")
print(t)

#compute the critical value at.05 significance level
alpha <- 0.01
t.alpha <- qt(1-alpha, df=n-1)
print("The value of t is")
print(-t.alpha)

if(t <= -t.alpha){
  print("reject the null hypothesis that the average life span of a battery is 2 or more years")
} else {
  print("Do not reject the nul;l hypothesis that the average life span of a battery is 2 or more years")
}
```

### Two-Tailed Test of Population Mean with Unknown Variance

> tTestBoth.R

```{r}
muO <- 15.4       # hypothesized value
xbar <- 14.6      # sample mean
s <- 2.5          # population standard deviation
n <- 29         # sample size

t <- (xbar-muO)/(s/sqrt(n))   # test statistic
print("t is equal to ")
print(t)

#compute the critical value at.05 significance level
alpha = 0.05
t.half.alpha <- qt(1-alpha/2, df=n-1)
print("t.half.alpha value is equal to")
print(-t.half.alpha, t.half.alpha)

if(t <= -t.half.alpha || t >= t.half.alpha){
  print("reject the null hypothesis that the mean penguin weight does not differ from last year")
} else {
  print("Do not reject the null hypothesis that the mean penguin weight does not differ from last year")
}
```

## Section 3: P-value

### PvalueEx1.R
> Solve the Lower Tail Test of Population Mean with Known Variance
> using p-value

```{r}
muO <- 10000 # hypothesized value
xbar <- 9900 # sample mean
sigma <- 120 # population standard deviation
n <- 30      # sample size
alpha <- .05

z <- (xbar-muO)/(sigma/sqrt(n)) # test statistic

print("z is equal to")
print(z)

# apply the pnorm function to compute the lower tail p-value of the test statistic
# As  it turns out to be less
# than the .05 significance level, we reject the null hypothesis that

pval <- pnorm(z)
print("pval is equal to")
print(pval)

if(pval < alpha)
  print("reject the null hypothesis")
```

### PvalueEx2.R
> Solve the Two-Tailed Test of Population Mean with Known Variance
> using p-value

```{r}
muO <-   15.4      # hypothesized value
xbar <-  14.6      # sample mean
sigma <-  2.5      # population standard deviation
n <-       35      # sample size
alpha <- 0.05

z <- (xbar-muO)/(s/sqrt(n))   # test statistic
print("z is equal to ")
print(z)

#apply the pnorm function to compute the two-tailed p-value of the test statistic.
#IT doubles the lower tailed p-value as the sample mean is less than the hypothesized value.
pval = 2 * pnorm(z) #lower tail
print("The value of pval is:")
print(pval)         # two-tailed p-value

if(pval < alpha){
  print("Reject the null hypothesis")
} else {
  print("Accept the null hypothesis")
}
```

### PvalueEx3.R
> Solve Lower Tail Test of Population Mean with Unknown Variance
> using p-value

```{r PValueEx3, message=TRUE, paged.print=TRUE}
muO <- 2     # hypothesized value
xbar <- 1.8  # sample mean
n <- 10      # population standard deviation
s <- 35      # sample size
alpha <- 0.01 

t <- (xbar-muO)/(s/sqrt(n))   
print("The value of t is equal to ")
print(t)

#apply the pt function to compute the lower tail p-value of the test statistic.
pval = pt(t, df=n-1)
print("Pval is equal to:")
print(pval)         # lower tail p-value

if(pval < alpha){
  print("Reject the null hypothesis")
} else {
  print("Accept the null hypothesis")
}
```

